### 2017-10-18 Draft2/Appending.Growth.tex
Rename some commands to PascalCase
More casing
Hard-code O and P

\groarrayimpl

\tcomplex

@@

\begin{align*}
k &= \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil\\
\bigo(k) &= \bigo(\log \varnitems)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\nwritesnewfn(\varnitems)$, leaving just $\biggo(\varnitems)$:

\begin{align*}
\biggo(\nwritesnewfn(\varnitems)) &= \biggo(\varnitems + \nwritesfn(k))\\
&= \biggo(\varnitems) + \biggo(\nwritesfn(k))\\
&= \biggo(\varnitems) + \bigo(\log \varnitems)\\
&= \biggo(\varnitems)
\end{align*}

\scomplex

@@

I determined in [...] that if $\varnitems$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $\bigo(\spacefn(k)) = \bigo(k) = \bigo(\log \varnitems)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\spacenewfn(\varnitems) = \biggo(2^{\lceil \log_2 \varnitems \rceil}) = \bigo(\varnitems)
\end{align*}

>>>>

\HdrGrowthArrayImpl

\HdrTimeComplex

@@

\begin{align*}
k &= \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil\\
O(k) &= O(\log \varnitems)
\end{align*}

Then, it is shown that the $k$-term disappears when analyzing the big-P complexity of $\nwritesnewfn(\varnitems)$, leaving just $P(\varnitems)$:

\begin{align*}
P(\nwritesnewfn(\varnitems)) &= P(\varnitems + \nwritesfn(k))\\
&= P(\varnitems) + P(\nwritesfn(k))\\
&= P(\varnitems) + O(\log \varnitems)\\
&= P(\varnitems)
\end{align*}

\HdrSpaceComplex

@@

I determined in [...] that if $\varnitems$ are appended then the tail is appended to $k$ times, where $k = \lceil \log_{\growthfactornew} \varnitems - \log_{\growthfactornew} \initcapacitynew \rceil$. Since the tail is a dynamic array, it allocates $\spacefn(k)$ space, where $\spacefn$ is the space function of the tail. Through a similar method to [...] it can be shown $O(\spacefn(k)) = O(k) = O(\log \varnitems)$. Then % TODO: P-analysis should come after the explicit formula.

\begin{align*}
\spacenewfn(\varnitems) = P(2^{\lceil \log_2 \varnitems \rceil}) = O(\varnitems)
\end{align*}

### 2017-10-18 Draft2/Appendix.tex
Hard-code O and P

\begin{theorem}
	$\biggo$ merges over addition and subtraction. That is,
	\begin{align*}
	\biggo(f(n)) + \biggo(g(n)) &= \biggo(f(n) + g(n))\\
	\biggo(f(n)) - \biggo(g(n)) &= \biggo(f(n) - g(n))
	\end{align*}
\end{theorem}

@@

	\textit{Let $f' \in \biggo(f(n))$ and $g' \in \biggo(g(n))$. Then $f' + g' \in \biggo(f(n) + g(n))$.}

	We can show this by using the limit definition of $\biggo$. For brevity, we will omit the parameter $n$ to each function. Since

@@

	it follows that $f' + g' \in \biggo(f(n) + g(n))$. The statement about subtraction can be proved in a similar fashion.
\end{proof}

\begin{theorem}
	$\biggo$ merges over multiplication and division. That is,
	\begin{align*}
	\biggo(f(n))\biggo(g(n)) &= \biggo(f(n)g(n))\\
	\frac {\biggo(f(n))} {\biggo(g(n))} &= \biggo\left(\frac{f(n)} {g(n)}\right)
	\end{align*}
\end{theorem}

@@

	\textit{Let $f' \in \biggo(f(n))$ and $g' \in \biggo(g(n))$. Then $f'g' \in \biggo(f(n)g(n))$.}

@@

	it follows that $f'g' \in \biggo(f(n)g(n))$. The statement about division can be proved by flipping the fraction for $g'$ before multiplying, resulting in

@@

	This implies $f' / g' \in \biggo(f(n) / g(n))$.
\end{proof}

@@

\begin{lemma}
	Each function $f(n)$ has a unique class of functions under $\biggo$. That is, suppose $f \in \biggo(g(n))$ and $f \in \biggo(h(n))$. Then $\biggo(g(n)) = \biggo(h(n))$.
\end{lemma}

@@

	\begin{align*}
	\left( \lim_{n \to \infty} \frac {g(n)} {f(n)} \right) \left( \lim_{n \to \infty} \frac {f(n)} {h(n)} \right) &= 1\\
	\lim_{n \to \infty} \frac {g(n)} {h(n)} &= 1\\
	g &\in \biggo(h(n))
	\end{align*}

	Now, consider any $g' \in \biggo(g(n))$. Since

@@

	it follows that $g' \in \biggo(h(n))$, and thus $\biggo(g(n)) \subseteq \biggo(h(n))$. By a similar argument, it can be shown $\biggo(h(n)) \subseteq \biggo(g(n))$. Thus it must be true that $\biggo(g(n)) = \biggo(h(n))$.
\end{proof}

@@

\begin{theorem}
	If $\lim_{n \to \infty} \frac{g(n)}{f(n)} = 0$, then $\biggo(f(n) + g(n)) = \biggo(f(n))$.
\end{theorem}

@@

	Then $f(n) + g(n) \in \biggo(f(n))$. Since it is also true that $f(n) + g(n) \in \biggo(f(n) + g(n))$, it follows from \ref{UniqueClassOfFunctions} that $\biggo(f(n)) = \biggo(f(n) + g(n))$.
\end{proof}

>>>>

\begin{theorem}
	$P$ merges over addition and subtraction. That is,
	\begin{align*}
	P(f(n)) + P(g(n)) &= P(f(n) + g(n))\\
	P(f(n)) - P(g(n)) &= P(f(n) - g(n))
	\end{align*}
\end{theorem}

@@

	\textit{Let $f' \in P(f(n))$ and $g' \in P(g(n))$. Then $f' + g' \in P(f(n) + g(n))$.}

	We can show this by using the limit definition of $P$. For brevity, we will omit the parameter $n$ to each function. Since

@@

	it follows that $f' + g' \in P(f(n) + g(n))$. The statement about subtraction can be proved in a similar fashion.
\end{proof}

\begin{theorem}
	$P$ merges over multiplication and division. That is,
	\begin{align*}
	P(f(n))P(g(n)) &= P(f(n)g(n))\\
	\frac {P(f(n))} {P(g(n))} &= P\left(\frac{f(n)} {g(n)}\right)
	\end{align*}
\end{theorem}

@@

	\textit{Let $f' \in P(f(n))$ and $g' \in P(g(n))$. Then $f'g' \in P(f(n)g(n))$.}

@@

	it follows that $f'g' \in P(f(n)g(n))$. The statement about division can be proved by flipping the fraction for $g'$ before multiplying, resulting in

@@

	This implies $f' / g' \in P(f(n) / g(n))$.
\end{proof}

@@

\begin{lemma}
	Each function $f(n)$ has a unique class of functions under $P$. That is, suppose $f \in P(g(n))$ and $f \in P(h(n))$. Then $P(g(n)) = P(h(n))$.
\end{lemma}

@@

	\begin{align*}
	\left( \lim_{n \to \infty} \frac {g(n)} {f(n)} \right) \left( \lim_{n \to \infty} \frac {f(n)} {h(n)} \right) &= 1\\
	\lim_{n \to \infty} \frac {g(n)} {h(n)} &= 1\\
	g &\in P(h(n))
	\end{align*}

	Now, consider any $g' \in P(g(n))$. Since

@@

	it follows that $g' \in P(h(n))$, and thus $P(g(n)) \subseteq P(h(n))$. By a similar argument, it can be shown $P(h(n)) \subseteq P(g(n))$. Thus it must be true that $P(g(n)) = P(h(n))$.
\end{proof}

@@

\begin{theorem}
	If $\lim_{n \to \infty} \frac{g(n)}{f(n)} = 0$, then $P(f(n) + g(n)) = P(f(n))$.
\end{theorem}

@@

	Then $f(n) + g(n) \in P(f(n))$. Since it is also true that $f(n) + g(n) \in P(f(n) + g(n))$, it follows from \ref{UniqueClassOfFunctions} that $P(f(n)) = P(f(n) + g(n))$.
\end{proof}

### 2017-10-18 Draft2/BigPNotation.tex
Add proofs for big-P properties
Hard-code O and P

In order to highlight the benefit of growth arrays, I will define a new notation for time complexity. The reason for this is, for certain operations, growth arrays are only better than dynamic arrays by a constant factor. For example, dynamic arrays might allocate roughly $2\varnitems$ memory for appending $\varnitems$ items, while growth arrays would allocate roughly $\varnitems$ memory. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $\bigo(\varnitems)$.

@@

I define an alternative to big-O notation, which I will call \textbf{big-P notation}. I define $\biggo(f(n))$ to be the class of functions $g$ such that $g(n) \to f(n)$ for large $n$. Formally,

$$
g \in \biggo(f(n)) \leftrightarrow \lim_{n \to \infty} \frac{g(n)}{f(n)} = 1
$$

Notice that while $\bigo(2\varnitems) = \bigo(\varnitems)$, $\biggo(2\varnitems) \neq \biggo(\varnitems)$. Thus, $\biggo$ makes it possible to distinguish between a function that uses $\varnitems$ space and one that uses $2\varnitems$ space. \textbf{Note:} A consequence of this is that bases for logarithms must be specified in big-P notation.

\subsection{Properties}

\textbf{Note:} Throughout this paper, when $\biggo(f(n))$ or $\bigo(f(n))$ is used in an arithmetic expression, I am referring to a function in the set and not the set itself. For example, the verbal equivalent of $\frac{\biggo(2n)}{\biggo(n)} = \biggo(2)$ states "the quotient of a function in $\biggo(2n)$ and a function in $\biggo(n)$ is equal to a function in $\biggo(2)$."

A useful property $\biggo$ shares with $\bigo$ is that it merges over arithmetic operations. This is stated by the following two theorems, which are proved in \ref{MergingOverArithmetic}.

\begin{theorem}
	$\biggo$ merges over addition and subtraction. That is,
	\begin{align*}
	\biggo(f(n)) + \biggo(g(n)) &= \biggo(f(n) + g(n))\\
	\biggo(f(n)) - \biggo(g(n)) &= \biggo(f(n) - g(n))
	\end{align*}
\end{theorem}

\begin{theorem}
	$\biggo$ merges over multiplication and division. That is,
	\begin{align*}
	\biggo(f(n)) \cdot \biggo(g(n)) &= \biggo(f(n) \cdot g(n))\\
	\frac {\biggo(f(n))} {\biggo(g(n))} &= \biggo\left(\frac{f(n)} {g(n)}\right)
	\end{align*}
\end{theorem}

Another nice property $\biggo$ and $\bigo$ have in common is that lower-order terms are removed from consideration. For example, $\biggo(n) = \biggo(n + \log_2 n)$. This is stated by the below theorem, which is proved in \ref{ScrubsLowerOrderTerms}.

\begin{theorem}
	If $\lim_{n \to \infty} \frac{g(n)}{f(n)} = 0$, then $\biggo(f(n) + g(n)) = \biggo(f(n))$.
\end{theorem}

>>>>

In order to highlight the benefit of growth arrays, I will define a new notation for time complexity. The reason for this is, for certain operations, growth arrays are only better than dynamic arrays by a constant factor. For example, dynamic arrays might allocate roughly $2\varnitems$ memory for appending $\varnitems$ items, while growth arrays would allocate roughly $\varnitems$ memory. Even though growth arrays are clearly better in this regard, the big-O space complexity for both data structures would be the same, $O(\varnitems)$.

@@

I define an alternative to big-O notation, which I will call \textbf{big-P notation}. I define $P(f(n))$ to be the class of functions $g$ such that $g(n) \to f(n)$ for large $n$. Formally,

$$
g \in P(f(n)) \leftrightarrow \lim_{n \to \infty} \frac{g(n)}{f(n)} = 1
$$

Notice that while $O(2\varnitems) = O(\varnitems)$, $P(2\varnitems) \neq P(\varnitems)$. Thus, $P$ makes it possible to distinguish between a function that uses $\varnitems$ space and one that uses $2\varnitems$ space. \textbf{Note:} A consequence of this is that bases for logarithms must be specified in big-P notation.

\subsection{Properties}
\label{BigPProperties}

\textbf{Note:} Throughout this paper, when $P(f(n))$ or $O(f(n))$ is used in an arithmetic expression, I am referring to a function in the set and not the set itself. For example, the verbal equivalent of $\frac{P(2n)}{P(n)} = P(2)$ states "the quotient of a function in $P(2n)$ and a function in $P(n)$ is equal to a function in $P(2)$."

A useful property $P$ shares with $O$ is that it merges over arithmetic operations. This is stated by the following two theorems, which are proved in \ref{MergingOverArithmetic}.

\begin{theorem}
	$P$ merges over addition and subtraction. That is,
	\begin{align*}
	P(f(n)) + P(g(n)) &= P(f(n) + g(n))\\
	P(f(n)) - P(g(n)) &= P(f(n) - g(n))
	\end{align*}
\end{theorem}

\begin{theorem}
	$P$ merges over multiplication and division. That is,
	\begin{align*}
	P(f(n))P(g(n)) &= P(f(n)g(n))\\
	\frac {P(f(n))} {P(g(n))} &= P\left(\frac{f(n)} {g(n)}\right)
	\end{align*}
\end{theorem}

Another nice property $P$ and $O$ have in common is that lower-order terms are removed from consideration. For example, $P(n) = P(n + \log_2 n)$. This is stated by the below theorem, which is proved in \ref{ScrubsLowerOrderTerms}.

\begin{theorem}
	If $\lim_{n \to \infty} \frac{g(n)}{f(n)} = 0$, then $P(f(n) + g(n)) = P(f(n))$.
\end{theorem}

### 2017-10-18 Draft2/CopyArray.Blurb.tex
Rename some commands to PascalCase
Hard-code O and P

\descriptn

@@

\begin{itemize}
	\item Plain arrays hold on to exactly the amount of memory needed to hold their elements. However, dynamic and growth arrays allocate more space than necessary to optimize appending new items.
	\item The user wants to call a function in third-party code that takes a plain array as an argument.
	\item
	\item Plain arrays are contiguous, while growth arrays are fragmented and have worse locality.
	\item The indexer of growth arrays is several times slower than that of plain arrays, whether the $\bigo(1)$ or $\bigo(\log \varnitems)$ implementation is chosen.
\end{itemize}

>>>>

\HdrDescription

@@

\begin{itemize}
	\item Plain arrays hold on to exactly the amount of memory needed to hold their elements. However, dynamic and growth arrays allocate more space than necessary to optimize appending new items.
	\item The user wants to call a function in third-party code that takes a plain array as an argument.
	\item
	\item Plain arrays are contiguous, while growth arrays are fragmented and have worse locality.
	\item The indexer of growth arrays is several times slower than that of plain arrays, whether the $O(1)$ or $O(\log \varnitems)$ implementation is chosen.
\end{itemize}
